|

### 2021-12-16 01-Chp-Background.Rmd
End of day:

It then introduces evidence synthesis as the key framework used to guide the research presented in the remaining chapters. Finally, it outlines the aims, objectives and structure of this thesis, and briefly summarises the contributions to the scientific literature that arose from this research.

@@

The remaining 10-30% of cases are caused by other dementia subtypes (e.g. Lewy body or frontotemporal dementias) or other neurological diseases (e.g. Parkinson's disease).[@burns2009]

@@

Many cognitive assessment tools exist for the purpose of informing a diagnoses of dementia,[@sheehan2012] with two of the best known of these being the Mini Mental State Exam (MMSE) and Montreal Cognitive Assessment (MoCA) scale. The distinction between these memory scales and diagnostic criteria presented above should be noted. For example, the MMSE is used to provide evidence for part A of the criteria presented in \@ref(tab:diagnosticCriteria-table). Taken alone, it does not indicate the absence or presence of dementia, instead merely indicating cognitive impairment which could have another cause (for example, temporary delirium as a result of an infection or surgery).

@@

<!----------------------------------------------------------------------------->
(ref:diagnosticCriteria-caption) __Overview of the DSM-5 criteria for dementia (major neurocognitive disorder).__ - These published criteria[@edition2013] are commonly used to define dementia cases. Criterion A captures the cognitive decline aspect of dementia, while Criteria B-D ensure that the impairment observed is sufficient to interfere with daily living and can not explained by other causes. Note that all criteria (A-D) must be met.

@@

The blood lipid profile contains a range of component parts, or fractions. However, this thesis will only consider the two most important fractions, trigylcerides (TG) and cholesterol, which are either absorbed from food (exogenous lipids) or produced internally (endogenous lipids).[@feingold2000]

Triglycerides are the simplest and most common type of lipids, used to store excess calories from food and to move energy around the body.[@laufs2020] In contrast, cholesterol is primarily used to create cell walls and certain sex hormones.[@zampelas2019] Lipids are not water soluble, and so cholesterol is transported through the blood stream in lipoprotein structures of varying densities. Low-density-lipoprotein cholesterol (LDL-c), commonly know as "bad" cholesterol, acts as an energy conveyor by transporting fat to cells. In contrast, high density-lipoprotein cholesterol (HDL-c) transports cholesterol to the liver to be broken down and excreted.[@feingold2000]

@@

\begin{equation}
  TC \approx LDLc + HDLc + kTG
  (\#eq:total-cholesterol-formula)
\end{equation}

where $k$ is 0.20 if measurements are in milligrams per decilitre (_mg/dl_) and 0.45 if measured in millimole per litre (_mmol/l_). Widely-used ranges for the acceptable levels of different types of lipids are based on the National Cholesterol Education Program (NCEP)[@national2002third], and are outlined in Table \@ref(tab:lipidLevels-table).

Elevated LDL-c in the bloodstream, a condition also known as hypercholesterolaemia or hyperlipidaemia,[@nelson2013] can lead to atherosclerosis,[@libby2019] the build-up of fatty deposits in the blood vessels. These deposits constrict blood flow and can lead to vascular complications. Alternatively, part of the deposit can detach from the artery walls, forming a clot that can lead to a heart attack or stroke.[@libby2019]

@@

Statins are a commonly prescribed method of lipid regulation.[@collins2016] Statins inhibit the conversion of 3-hydroxy-3-methylglutaryl-coenzyme-A (HMG-CoA) into mevalonate, by competitively binding with HMG-CoA reductase (HMGCR). This conversion is a key rate-limiting step in the cholesterol biosynthesis pathway (see Figure \@ref(fig:statin-mechanisam)), enabling statins to reduce effectively the production of LDL cholesterol.

@@

(ref:statin-mechanisam-cap) __Statin mechanism of action__ - Statins inhibit HMG-CoA reductase which controls the conversion of HMG-CoA to mevalonate. This conversion is the rate-limiting step in cholesterol biosynthesis, and so statin use can lower cholesterol levels.

@@

Several statin treatments have been widely available for some time (see Table \@ref(tab:statinOverview-table)). Depending on the statin and dosage prescribed, the average reduction in LDL-c concentrations ranges from 15% with low-intensity regimen (e.g. ravastatin 5 mg/day) up to 60% with a high-intensity regimen (e.g. rosuvastatin 80 mg/day).[@collins2016; @law2003] Statins also vary with regard to their lipophilicity (the extent to which they are lipid soluble), affecting their localisation within the body, with hydophilic statins being concentrated in the liver and lipophilic statins circulating more widely.[@schachter2005] This may create a divide in the off-target effects of statins with differing lipophilicity, particularly given the ability of lipophilic statins to permeate the blood brain barrier.[@sierra2011]

@@

(ref:statinOverview-caption) __Overview of common statins__ -  There are several commonly-prescribed statins used to treat hypercholesterolemia. Their approval date (US), properties and lipid-lowering effect are described here.

@@

There are several other interventions that can be used to modify a persons lipid profile, each of which act via different mechanisms. However, in general, these treatments are either used as adjunct (additional) treatments to statin therapy or are used in situations where statins are contra-indicated or not tolerated.

@@

Fibrates provide a second example of non-statin therapy. They are used to treat hypertriglyceridaemia by reducing production of triglyceride carrying compounds in the liver. They are commonly used in patients with mixed hyperlipidaemia if treatment with statins has failed to sufficiently control cholesterol levels.

@@

Other therapies targeting triglycerides exist, including nicotinic acids[@mckenney2004] and omega-3-fatty acids,[@skulas2019] but they far less effective in LDL-c lowering than the therapies described above.

@@

Newer methodological approaches, such as Mendelian randomisation (MR),[@daveysmith2014] have also been used to examine the effect of varying lipid levels on dementia risk. Mendelian randomisation attempts to combat the risk of reverse causation and residual confounding inherent to observational studies by using natural random varation in participants' genomes.[@greenland2000] In brief, MR uses genetic variants that are both strongly associated with the exposure of interest and are independent from potential confounders to strengthen causal inference.[@daveysmith2014] The analytic method relies on several assumptions about the instrumental variable (IV)[@davies2018] which are discussed in Section \@ref(rev-discussion-MR).

@@

One of the major criticisms of using preprints as an evidence source is that they have not yet undergone formal peer review.[@maslove2018; @schalkwyk2020] However, this approach assigns substantial weight to peer-review as a indicator of "quality", and is at odds with the acceptance of non-reviewed conference proceedings as an evidence source.[@lefebvre2019searching; @mahood2014] The argument for including preprints as an evidence source is further strengthened by results that demonstrate preprinted studies seldom change following peer review. Meta-studies of the concordance between preprinted and published studies showed that results were broadly comparable between the two, indicating that while the numerical results may change, the overall interpretation of the results were consistent in the majority of cases.[@shi2021; @klein2019; @nicholson2021] This indicates that preprints should be considered a reliable reflection of a given study.

@@

I will attempt to obtain the raw data from relevant primary studies identified by the systematic review presentd in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading). Any data obtained will be combined with that available from the DPUK portal as part of an IPD meta-analysis in Chapter \@ref(ipd-heading). This will enable the assessment of the effect of lipids on dementia, and whether this relationship is modified by key variables such as participants' age or sex.

@@

As illustrated in Section \@ref(evidence-association), several diverse epidemiological methods have been used to examine the effect of varying blood lipid levels on dementia risk. Aetiological triangulation is a developing evidence synthesis method that seeks to exploit these differences in study design, and as a result, in underlying bias structures.[@lawlor2016] If several sources of evidence are available and point towards identical conclusions about an exposure-outcome relationship, and these sources are at risk of unrelated biases, this strengthens our confidence in the result. The ideal scenario is where predicted sources of bias are likely to be in competing directions, one to strengthen the effect of the exposure and the other to attenuate it.[@lawlor2016] As such, triangulating these results can provide a middle-ground between the competing directions of bias. A triangulation approach can also prove useful in helping to prospectively design new studies that are at risk of different sources of bias to those already observed in the published literature.[@munafo2018] However, existing methods for triangulation are limited to a qualitative discussion of the different identified results with respect to their biases,[@lawlor2016] an approach which faces issues in interpretability when assessing many results.

@@

* To create a tool that allows for the inclusion of health related preprints in evidence syntheses in a systematic and reproducible manner
* To review all available evidence across multiple diverse study designs to assess the effect of lipids and lipid regulating agents on dementia risk
* To examine whether there is evidence for an effect of lipid-regulating agents on dementia and related outcomes in a large scale population-based cohort, the Clinical Practice Research Datalink (CPRD)
* To produce evidence on lipid-covariate interactions as part of an IPD meta-analysis 
* To propose a generalised framework for the quantitative triangulation across diverse evidence sources

@@

-   **Chapter \@ref(background-heading):** Background information on dementia and blood lipid levels. This chapter provides an introduction to the topics covered in this thesis, and discusses the motivation for the research presented in the following chapters.
-   **Chapter \@ref(sys-rev-tools-heading):** This chapter introduces a new tool, `medrxivr`, which was developed to allow for systematic searches of the health-related preprint repositories.
-   **Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading):** These chapters describe, respectively, the methods and results of a comprehensive systematic review and meta-analysis. This review examined all available evidence on the association of blood lipids and LRA with subsequent risk of dementia.
-   **Chapter \@ref(cprd-analysis-heading):** This chapter examines the relationship between lipid-regulating agent use and dementia outcomes in the Clinical Practice Research Datalink, a large primary care electronic health record database.
-   **Chapter \@ref(ipd-heading):** This chapter describes an individual participant data analysis of three longitudinal cohort studies to describe the relationship between blood lipids and dementia outcomes
-   **Chapter \@ref(tri-heading)**: This chapter integrates the evidence identified and produced by the preceding chapters as part of a quantitative triangulation framework.
-   **Chapter \@ref(discussion-heading)**: This chapter summarise the key clinical and methodological findings of the thesis and discusses their implications for practice. The overall strengths and weaknesses of this project are discussed in detail, and further avenues of research are suggested.

>>>>

It then introduces evidence synthesis as the key framework used to guide the research presented in the remaining chapters. Finally, it outlines the aims, objectives, and structure of this thesis, and briefly summarises the contributions to the scientific literature that arose from this research.

@@

The remaining 10-30% of cases are caused by other dementia subtypes (e.g., Lewy body or frontotemporal dementias) or other neurological diseases (e.g., Parkinson's disease).[@burns2009]

@@

Many cognitive assessment tools exist for the purpose of informing a diagnosis of dementia,[@sheehan2012] with two of the best known of these being the Mini Mental State Exam (MMSE) and Montreal Cognitive Assessment (MoCA) scale. The distinction between these memory scales and diagnostic criteria presented above should be noted. For example, the MMSE is used to provide evidence for part A of the criteria presented in \@ref(tab:diagnosticCriteria-table). Taken alone, it does not indicate the absence or presence of dementia, instead merely indicating cognitive impairment which could have another cause (for example, temporary delirium as a result of an infection or surgery).

@@

<!----------------------------------------------------------------------------->
(ref:diagnosticCriteria-caption) __Overview of the DSM-5 criteria for dementia (major neurocognitive disorder).__ - These published criteria[@edition2013] are commonly used to define dementia cases. Criterion A captures the cognitive decline aspect of dementia, while Criteria B-D ensure that the impairment observed is sufficient to interfere with daily living and cannot explained by other causes. Note that all criteria (A-D) must be met.

@@

The blood lipid profile contains a range of component parts, or fractions. However, this thesis will only consider the two most important fractions, triglycerides (TG) and cholesterol, which are either absorbed from food (exogenous lipids) or produced internally (endogenous lipids).[@feingold2000]

Triglycerides are the simplest and most common type of lipids, used to store excess calories from food and to move energy around the body.[@laufs2020] In contrast, cholesterol is primarily used to create cell walls and certain sex hormones.[@zampelas2019] Lipids are not water soluble, and so cholesterol is transported through the blood stream in lipoprotein structures of varying densities. Low-density-lipoprotein cholesterol (LDL-c), commonly known as "bad" cholesterol, acts as an energy conveyor by transporting fat to cells. In contrast, high density-lipoprotein cholesterol (HDL-c) transports cholesterol to the liver to be broken down and excreted.[@feingold2000]

@@

\begin{equation}
  TC \approx LDL + HDL + kTG
  (\#eq:total-cholesterol-formula)
\end{equation}

where $k$ is 0.20 if measurements are in milligrams per decilitre (_mg/dl_) and 0.45 if measured in millimole per litre (_mmol/l_). Widely used ranges for the acceptable levels of different types of lipids are based on the National Cholesterol Education Program (NCEP)[@national2002third], and are outlined in Table \@ref(tab:lipidLevels-table).

Elevated LDL-c in the bloodstream, a condition also known as hypercholesterolemia or hyperlipidaemia,[@nelson2013] can lead to atherosclerosis,[@libby2019] the build-up of fatty deposits in the blood vessels. These deposits constrict blood flow and can lead to vascular complications. Alternatively, part of the deposit can detach from the artery walls, forming a clot that can lead to a heart attack or stroke.[@libby2019]

@@

Statins are a commonly prescribed method of lipid regulation.[@collins2016] Statins inhibit the conversion of 3-hydroxy-3-methylglutaryl-coenzyme-A (HMG-CoA) into mevalonate, by competitively binding with HMG-CoA reductase (HMGCR). This conversion is a key rate-limiting step in the cholesterol biosynthesis pathway (see Figure \@ref(fig:statin-mechanisam)), enabling statins to effectively reduce the production of LDL cholesterol.

@@

(ref:statin-mechanisam-cap) __Statin mechanism of action__ - Statins inhibit HMG-CoA reductase which controls the conversion of HMG-CoA to mevalonate. This conversion is the rate-limiting step in cholesterol biosynthesis, and so statin use can reduce the production of cholesterol.

@@

Several statin treatments have been widely available for some time (see Table \@ref(tab:statinOverview-table)). Depending on the statin and dosage prescribed, the average reduction in LDL-c concentrations ranges from 15% with low-intensity regimen (e.g., pravastatin 5 mg/day) up to 60% with a high-intensity regimen (e.g., rosuvastatin 80 mg/day).[@collins2016; @law2003] Statins also vary with regard to their lipophilicity (the extent to which they are lipid soluble), affecting their localisation within the body, with hydrophilic statins being concentrated in the liver and lipophilic statins circulating more widely.[@schachter2005] This may create a divide in the off-target effects of statins with differing lipophilicity, particularly given the ability of lipophilic statins to permeate the blood brain barrier.[@sierra2011]

@@

(ref:statinOverview-caption) __Overview of common statins__ -  There are several commonly prescribed statins used to treat hypercholesterolemia. Their approval date (US), properties and lipid-lowering effect are described here.

@@

There are several other interventions that can be used to modify a person's lipid profile, each of which act via different mechanisms. However, in general, these treatments are either used as adjunct (additional) treatments to statin therapy or are used in situations where statins are contra-indicated or not tolerated.

@@

Fibrates provide a second example of non-statin therapy. They are used to treat hypertriglyceridemia by reducing production of triglyceride carrying compounds in the liver. They are commonly used in patients with mixed hyperlipidaemia if treatment with statins has failed to sufficiently control cholesterol levels.

@@

Other therapies targeting triglycerides exist, including nicotinic acids[@mckenney2004] and omega-3-fatty acids,[@skulas2019] but they are far less effective in LDL-c lowering than the therapies described above.

@@

Newer methodological approaches, such as Mendelian randomisation (MR),[@daveysmith2014] have also been used to examine the effect of varying lipid levels on dementia risk. Mendelian randomisation attempts to combat the risk of reverse causation and residual confounding inherent to observational studies by using natural random variation in participants' genomes.[@greenland2000] In brief, MR uses genetic variants that are both strongly associated with the exposure of interest and are independent from potential confounders to strengthen causal inference.[@daveysmith2014] The analytic method relies on several assumptions about the instrumental variable (IV)[@davies2018] which are discussed in Section \@ref(rev-discussion-MR).

@@

One of the major criticisms of using preprints as an evidence source is that they have not yet undergone formal peer review.[@maslove2018; @schalkwyk2020] However, this approach assigns substantial weight to peer-review as an indicator of "quality", and is at odds with the acceptance of non-reviewed conference proceedings as an evidence source.[@lefebvre2019searching; @mahood2014] The argument for including preprints as an evidence source is further strengthened by results that demonstrate preprinted studies seldom change following peer review. Meta-studies of the concordance between preprinted and published studies showed that results were broadly comparable between the two, indicating that while the numerical results may change, the overall interpretation of the results were consistent in the majority of cases.[@shi2021; @klein2019; @nicholson2021] This indicates that preprints should be considered a reliable reflection of a given study.

@@

I will attempt to obtain the raw data from relevant primary studies identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading). Any data obtained will be combined with that available from the DPUK portal as part of an IPD meta-analysis in Chapter \@ref(ipd-heading). This will enable the assessment of the effect of lipids on dementia, and whether this relationship is modified by key variables such as participants' age or sex.

@@

As illustrated in Section \@ref(evidence-association), several diverse epidemiological methods have been used to examine the effect of varying blood lipid levels on dementia risk. Aetiological triangulation is a developing evidence synthesis method that seeks to exploit these differences in study design, and as a result, in underlying bias structures.[@lawlor2016] If several sources of evidence are available and point towards identical conclusions about an exposure-outcome relationship, and these sources are at risk of unrelated biases, this strengthens our confidence in the result. The ideal scenario is where predicted sources of bias are likely to be in competing directions, one to strengthen the effect of the exposure and the other to attenuate it.[@lawlor2016] As such, triangulating these results can provide a middle ground between the competing directions of bias. A triangulation approach can also prove useful in helping to prospectively design new studies that are at risk of different sources of bias to those already observed in the published literature.[@munafo2018] However, existing methods for triangulation are limited to a qualitative discussion of the different identified results with respect to their biases,[@lawlor2016] an approach which faces issues in interpretability when assessing many results.

@@

* To create a tool that allows for the inclusion of health-related preprints in evidence syntheses in a systematic and reproducible manner
* To review all available evidence across multiple diverse study designs to assess the effect of lipids and lipid regulating agents on dementia risk
* To examine whether there is evidence for an effect of lipid-regulating agents on dementia and related outcomes in a large-scale population-based cohort, the Clinical Practice Research Datalink (CPRD)
* To produce evidence on lipid-covariate interactions as part of an IPD meta-analysis 
* To propose a generalised framework for the quantitative triangulation across diverse evidence sources

@@

-   **Chapter \@ref(background-heading):** Background information on dementia and blood lipid levels. This chapter provides an introduction to the topics covered in this thesis and discusses the motivation for the research presented in the following chapters.
-   **Chapter \@ref(sys-rev-tools-heading):** This chapter introduces a new tool, `medrxivr`, which was developed to allow for systematic searches of the health-related preprint repositories.
-   **Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading):** These chapters describe, respectively, the methods and results of a comprehensive systematic review and meta-analysis. This review examined all available evidence on the association of blood lipids and LRA with subsequent risk of dementia.
-   **Chapter \@ref(cprd-analysis-heading):** This chapter examines the relationship between lipid-regulating agent use and dementia outcomes in the Clinical Practice Research Datalink, a large primary care electronic health record database.
-   **Chapter \@ref(ipd-heading):** This chapter describes an individual participant data analysis of three longitudinal cohort studies to describe the relationship between blood lipids and dementia outcomes
-   **Chapter \@ref(tri-heading)**: This chapter integrates the evidence identified and produced by the preceding chapters as part of a quantitative triangulation framework.
-   **Chapter \@ref(discussion-heading)**: This chapter summarise the key clinical and methodological findings of the thesis and discusses their implications for practice. The overall strengths and weaknesses of this project are discussed in detail, and further avenues of research are suggested.

### 2021-12-16 02-Chp-Systematic-Review-Tools.Rmd
Range of updates
End of day:

2.  support for bulk export of references returned to a file type that can be readily imported into a reference manager (e.g., *.bib* or *.ris*); and

@@

Prior to searching the metadata, users must first create a local copy of the database, denoted using `mx_data` below. In `medrixvr`, I have provided two separate but related methods to achieve this (Figure \@ref(fig:medrxivr-data-sources)). The first of these methods, accessed via the `mx_api_content()` function, creates a local copy of all data available from the medRxiv API at the time the function is run.

@@

```{r, eval = FALSE, class.source = "bg-danger"}
# Get a copy of the database from the live medRxiv API endpoint
mx_data <- mx_api_content()
```

@@

This provides an up-to-the-minute reflection of the medRxiv preprint repository. However, this approach has two limitations. Firstly, because the API returns results as a series of pages limited to 100 records per page, downloading the entire database requires a time-intensive process of cycling through each page. Secondly, the API can become unavailable, either during peak usage times or planned maintainence windows. To address these limitations, I provide a second method of accessing medRxiv data, called via the `mx_snapshot()` function, which allows users to access a maintained static snapshot of the database. Note that due to the size of bioRxiv, only a maintained snapshot of the medRxiv repository is available via `mx_snapshot()`.

@@

For example, when written in standard syntax, the search contained in the `myquery` object above would be: "((dementia **OR** alzheimer's) **AND** (lipids **OR** statins))". There is no limit to the number of search terms that can be included in each topic, nor in the number of topics that can be searched for. Search terms can also contain common syntax used by systematic reviewers and health librarians, including the NEAR operator for the identification of co-localised terms and wild-cards which allow for alternate spellings (e.g. "randomi*s*ation" vs "randomi*z*ation").

@@

This is important as part of developing a search strategy,[@bramer2018] because it allows for the key terms related to each topic to be discovered. It also aids in identifying misspelled or case-sensitive search terms, which will frequently return no results. As an example, in the search presented above, the term "alzheimer's" returns no records. This is expected, because "Alzheimer's" is a proper noun and so should be capitalised, but serves to illustrate the usefulness of the reporting function.

@@

In line with my second success criteria (Section \@ref(success-criteria)), one of the key features of the `medrxivr` is the ability for users to export easily the results of their systematic search to a reference manager. While this would seem a likely candidate to be a generic feature, this is one of the key ways in which `medrxivr` is set apart from other preprint search tools. None of tools that I audited provide this feature, nor does the native medRxiv/bioRxiv website search.

The results of our simple search above can be exported to the `medrxiv_export.bib"` file using the following code:

@@

For example, I led a descriptive analysis of the change in data availability statements between preprinted and published versions of the same manuscript, stratified by journal data sharing policy access, underpinned by preprint meta-data provided by `medrixvr`. By comparing the preprinted and published versions of the data availability statement, I could examine the same manuscript - same content, authors and funders - under two different publication policies (preprint server vs peer-review journal). This comparison provides evidence on whether stricter policies which require data sharing as a condition of publication result in increased data availability. The analysis found some evidence that data availability statements more frequently described open data on publication when the journal mandated data sharing, compared to when the journal did not mandate data sharing. This study has since been published,[@mcguinness2020DAScomparison] and a copy is included in Appendix \@ref(published-papers).

@@

Secondly, using `medrxivr`, an analysis of the publication rate for medRxiv preprints was performed (see Appendix \@ref(appendix-tools-pub)). `r published` of the `r publication_n` records `r publish_percent` posted on medRxiv in July 2019 were published by 30th July 2021 (i.e. allowing for a two-year lag between preprint posting and publication). This finding agrees with previous work demonstrating that two-thirds of bioRxiv preprints are published in a peer-reviewed journal within two years of posting.[@abdill2019popularity] This analysis further illustrates that a substantial number of studies are never formally published but remain accessible as preprints.

In summary, the meta-research/methodological analyses described above illustrate that easy access to medRxiv/bioRxiv metadata has applications beyond the inclusion of preprints in systeamtic review.

@@

### Role of open source tools in evidence synthesis

Part of the motivation for creating the `medrxivr` tool was a belief that the development and distribution of open source scripts and tools should be a fundamental part of evidence synthesis research.[@goldacre2019; @mckiernan2016] In the case of `medrxivr`, it is likely that several other evidence synthesists had written personal scripts that have a similar, or related, functionality - in fact, following development of the tool, I identified one other researcher that has done so (Nicholas Fraser, author of the `rbiorxiv` package, which allows for importing medRxiv metadata into R but does not provide search functionality).[@fraser2020rbiorixv] If these scripts continue to be developed in private and are never shared or publicised, this will inevitably hamper the efforts of the evidence synthesis community, not only in terms of duplication of time and effort but also due to lost opportunities for collaboration.[@mckiernan2016] Creating and sharing well-documented packages, the recognised standard for sharing code in R, represents one way to reduce this inefficiency.[@vuorre2020]

>>>>

2.  support for bulk export of references returned to a file type that can be readily imported into a reference manager (e.g.,, *.bib* or *.ris*); and

@@

Prior to searching the metadata, users must first create a local copy of the database, denoted using `mx_data` below. In `medrxivr`, I have provided two separate but related methods to achieve this (Figure \@ref(fig:medrxivr-data-sources)). The first of these methods, accessed via the `mx_api_content()` function, creates a local copy of all data available from the medRxiv API at the time the function is run.

@@

```{r, eval = FALSE}
# Get a copy of the database from the live medRxiv API endpoint
mx_data <- mx_api_content()
```

@@

This provides an up-to-the-minute reflection of the medRxiv preprint repository. However, this approach has two limitations. Firstly, because the API returns results as a series of pages limited to 100 records per page, downloading the entire database requires a time-intensive process of cycling through each page. Secondly, the API can become unavailable, either during peak usage times or planned maintenance windows. To address these limitations, I provide a second method of accessing medRxiv data, called via the `mx_snapshot()` function, which allows users to access a maintained static snapshot of the database. Note that due to the size of bioRxiv, only a maintained snapshot of the medRxiv repository is available via `mx_snapshot()`.

@@

For example, when written in standard syntax, the search contained in the `myquery` object above would be: "((dementia **OR** alzheimer's) **AND** (lipids **OR** statins))". There is no limit to the number of search terms that can be included in each topic, nor in the number of topics that can be searched for. Search terms can also contain common syntax used by systematic reviewers and health librarians, including the NEAR operator for the identification of co-localised terms and wild cards which allow for alternate spellings (e.g., "randomi*s*ation" vs "randomi*z*ation").

@@

This is important as part of developing a search strategy,[@bramer2018] because it allows for the key terms related to each topic to be discovered. It also aids in identifying misspelled or case-sensitive search terms, which will frequently return no results. As an example, in the search presented above, the term "alzheimer's" returns no records. This is expected because "Alzheimer's" is a proper noun and so should be capitalised, but serves to illustrate the usefulness of the reporting function.

@@

In line with my second success criteria (Section \@ref(success-criteria)), one of the key features of the `medrxivr` is the ability for users to easily export the results of their systematic search to a reference manager. While this would seem a likely candidate to be a generic feature, this is one of the key ways in which `medrxivr` is set apart from other preprint search tools. None of tools that I audited provide this feature, nor does the native medRxiv/bioRxiv website search.

The results of our simple search above can be exported to the `medrxiv_export.bib` file using the following code:

@@

For example, I led a descriptive analysis of the change in data availability statements between preprinted and published versions of the same manuscript, stratified by journal data sharing policy access, underpinned by preprint meta-data provided by `medrxivr`. By comparing the preprinted and published versions of the data availability statement, I could examine the same manuscript - same content, authors and funders - under two different publication policies (preprint server vs peer-review journal). This comparison provides evidence on whether stricter policies which require data sharing as a condition of publication result in increased data availability. The analysis found some evidence that data availability statements more frequently described open data on publication when the journal mandated data sharing, compared to when the journal did not mandate data sharing. This study has since been published,[@mcguinness2020DAScomparison] and a copy is included in Appendix \@ref(published-papers).

@@

Secondly, using `medrxivr`, an analysis of the publication rate for medRxiv preprints was performed (see Appendix \@ref(appendix-tools-pub)). `r published` of the `r publication_n` records `r publish_percent` posted on medRxiv in July 2019 were published by 30th July 2021 (i.e., allowing for a two-year lag between preprint posting and publication). This finding agrees with previous work demonstrating that two-thirds of bioRxiv preprints are published in a peer-reviewed journal within two years of posting.[@abdill2019popularity] This analysis further illustrates that a substantial number of studies are never formally published but remain accessible as preprints.

In summary, the meta-research/methodological analyses described above illustrate that easy access to medRxiv/bioRxiv metadata has applications beyond the inclusion of preprints in systematic review.

@@

### Role of open-source tools in evidence synthesis

Part of the motivation for creating the `medrxivr` tool was a belief that the development and distribution of open-source scripts and tools should be a fundamental part of evidence synthesis research.[@goldacre2019; @mckiernan2016] In the case of `medrxivr`, it is likely that several other evidence synthesists had written personal scripts that have a similar, or related, functionality - in fact, following development of the tool, I identified one other researcher that has done so (Nicholas Fraser, author of the `rbiorxiv` package, which allows for importing medRxiv metadata into R but does not provide search functionality).[@fraser2020rbiorixv] If these scripts continue to be developed in private and are never shared or publicised, this will inevitably hamper the efforts of the evidence synthesis community, not only in terms of duplication of time and effort but also due to lost opportunities for collaboration.[@mckiernan2016] Creating and sharing well-documented packages, the recognised standard for sharing code in R, represents one way to reduce this inefficiency.[@vuorre2020]
